---
title: "CSB Project"
author: "Sharanya Sheshadri"
date: "`r Sys.Date()`"
output: html_document
---

# Load Packages
```{r, warning=FALSE, message=FALSE, results='hide'}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

# Define the packages we need
packages <- c("GEOquery", "limma", "hgu133a.db", "annotate", "clusterProfiler", "org.Hs.eg.db", "STRINGdb", "RCy3", "multiMiR", "TCGAbiolinks", "SummarizedExperiment")

# For each package, check if it's installed and install if not
for (pkg in packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    message(paste0("Installing package: ", pkg))
    BiocManager::install(pkg)
  } else {
    message(paste0("Package already installed: ", pkg))
  }
}

# Load all packages
lapply(packages, library, character.only = TRUE)

# Define the packages we need
packages <- c("WGCNA", "ggplot2", "dplyr", "stringr", "pheatmap", "ggVennDiagram", "igraph", "tidygraph", "ggraph", "visNetwork", "htmlwidgets", "webshot", "pdftools", "gridExtra", "grid", "ggrepel","survival", "survminer", "readr", "knitr", "kableExtra")

# For each package, check if it's installed and install if not
for (pkg in packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    message(paste0("Installing package: ", pkg))
    install.packages(pkg)
  } else {
    message(paste0("Package already installed: ", pkg))
  }
}

# Load all packages
apply(packages, library, character.only = TRUE)
```

# Download Data
```{r}
# Download the GEO datasets
gse41258 <- getGEO("GSE41258", GSEMatrix = TRUE)
gse68204 <- getGEO("GSE68204", GSEMatrix = TRUE)
gse62322 <- getGEO("GSE62322", GSEMatrix = TRUE)

# Extract expression matrices
expr_gse41258 <- exprs(gse41258[[1]])
expr_gse68204 <- exprs(gse68204[[1]])
expr_gse62322 <- exprs(gse62322[[1]])

# Extract phenotype data
pheno_gse41258 <- pData(gse41258[[1]])
pheno_gse68204 <- pData(gse68204[[1]])
pheno_gse62322 <- pData(gse62322[[1]])

# Extract feature data for mapping probes to genes
feat_gse41258 <- fData(gse41258[[1]])
feat_gse68204 <- fData(gse68204[[1]])
feat_gse62322 <- fData(gse62322[[1]])
```

# Preprocess Data
## GSE41258
```{r}
# 1. Filter GSE41258 to select tumor and normal samples
# Extract sample information to identify normal and tumor samples
sample_info_41258 <- pheno_gse41258$characteristics_ch1
tumor_indices <- grep("Primary Tumor", sample_info_41258)
normal_indices <- grep("Normal Colon", sample_info_41258)

# Extract tumor and normal samples
tumor_samples_41258 <- expr_gse41258[, tumor_indices]
normal_samples_41258 <- expr_gse41258[, normal_indices]

# Combine for DEG analysis later
combined_samples_41258 <- cbind(normal_samples_41258, tumor_samples_41258)
sample_type_41258 <- c(rep("normal", ncol(normal_samples_41258)), 
                       rep("tumor", ncol(tumor_samples_41258)))

# 2. Normalize the data using quantile normalization
combined_samples_41258_norm <- normalizeQuantiles(combined_samples_41258)

# 3. Check if the data is log2 transformed
# If the values are not in log2 scale (e.g., if they have a wide range like 0-10000),
# then transform them
if(max(combined_samples_41258_norm, na.rm = TRUE) > 100) {
  # Add a small value to avoid log(0)
  combined_samples_41258_norm <- log2(combined_samples_41258_norm + 1)
  print("Applied log2 transformation to GSE41258 data")
} else {
  print("GSE41258 data appears to be already log-transformed")
}

boxplot(combined_samples_41258_norm, outline=FALSE)
```

## GSE62322
```{r}
# GSE62322 (validation dataset)
# Filter samples
sample_info_62322 <- pheno_gse62322$characteristics_ch1
tumor_indices_62322 <- grep("Primary Tumor", sample_info_62322)
normal_indices_62322 <- grep("Normal Colon", sample_info_62322)

tumor_samples_62322 <- expr_gse62322[, tumor_indices_62322]
normal_samples_62322 <- expr_gse62322[, normal_indices_62322]

combined_samples_62322 <- cbind(normal_samples_62322, tumor_samples_62322)
sample_type_62322 <- c(rep("normal", ncol(normal_samples_62322)), 
                       rep("tumor", ncol(tumor_samples_62322)))

# Normalize
combined_samples_62322_norm <- normalizeQuantiles(combined_samples_62322)

# Check and transform if needed
range(combined_samples_62322_norm)
if(max(combined_samples_62322_norm, na.rm = TRUE) > 100) {
  combined_samples_62322_norm <- log2(combined_samples_62322_norm + 1)
  print("Applied log2 transformation to GSE62322 data")
} else {
  print("GSE62322 data appears to be already log-transformed")
}
```

## GSE68204
```{r}
# GSE68204 (miRNA dataset)
# Filter samples
sample_info_68204 <- pheno_gse68204$characteristics_ch1.2
tumor_indices_68204 <- grep("Tumour", sample_info_68204)
normal_indices_68204 <- grep("Normal", sample_info_68204)

tumor_samples_68204 <- expr_gse68204[, tumor_indices_68204]
normal_samples_68204 <- expr_gse68204[, normal_indices_68204]

combined_samples_68204 <- cbind(normal_samples_68204, tumor_samples_68204)
sample_type_68204 <- c(rep("normal", ncol(normal_samples_68204)), 
                       rep("tumor", ncol(tumor_samples_68204)))

# Normalize
combined_samples_68204_norm <- normalizeQuantiles(combined_samples_68204)

# Check and transform if needed
range(combined_samples_68204_norm)
if(max(combined_samples_68204_norm, na.rm = TRUE) > 100) {
  combined_samples_68204_norm <- log2(combined_samples_68204_norm + 1)
  print("Applied log2 transformation to GSE68204 data")
} else {
  print("GSE68204 data appears to be already log-transformed")
}
```

# Convert Probe ID to Genes
## GSE41258
```{r}
# Get the platform annotation
platform_gse41258 <- annotation(gse41258[[1]])
print(platform_gse41258)  # Should print "GPL96"

# Convert probe IDs to gene symbols for GSE41258
probe_ids_41258 <- rownames(combined_samples_41258_norm)
gene_symbols_41258 <- mapIds(hgu133a.db, 
                             keys = probe_ids_41258,
                             column = "SYMBOL",
                             keytype = "PROBEID",
                             multiVals = "first")

# Some probes might not map to gene symbols, handle those
genes_41258 <- data.frame(
  probe_id = probe_ids_41258,
  gene_symbol = gene_symbols_41258,
  stringsAsFactors = FALSE
)

# Check how many probes were mapped to gene symbols
sum(!is.na(genes_41258$gene_symbol))
sum(is.na(genes_41258$gene_symbol))

# Create a new expression matrix with gene symbols as row names
# For probes that map to the same gene, we'll keep the one with highest variance
expr_with_genes_41258 <- data.frame(combined_samples_41258_norm)
expr_with_genes_41258$gene_symbol <- gene_symbols_41258

# Remove probes without gene symbols
expr_with_genes_41258 <- expr_with_genes_41258[!is.na(expr_with_genes_41258$gene_symbol), ]

# For probes mapping to the same gene, keep the one with highest variance
expr_with_genes_41258$variance <- apply(expr_with_genes_41258[, 1:(ncol(expr_with_genes_41258)-2)], 1, var)
expr_with_genes_41258 <- expr_with_genes_41258[order(expr_with_genes_41258$gene_symbol, 
                                                    -expr_with_genes_41258$variance), ]
expr_with_genes_41258 <- expr_with_genes_41258[!duplicated(expr_with_genes_41258$gene_symbol), ]

# Create final expression matrix with gene symbols as row names
expr_gse41258_genes <- expr_with_genes_41258[, 1:(ncol(expr_with_genes_41258)-2)]
rownames(expr_gse41258_genes) <- expr_with_genes_41258$gene_symbol

```

## GSE62322
```{r}
# Get the platform annotation
platform_gse62322 <- annotation(gse62322[[1]])
print(platform_gse62322)  # Should print "GPL96"

# Convert probe IDs to gene symbols for GSE62322
probe_ids_62322 <- rownames(combined_samples_62322_norm)
gene_symbols_62322 <- mapIds(hgu133a.db, 
                             keys = probe_ids_62322,
                             column = "SYMBOL",
                             keytype = "PROBEID",
                             multiVals = "first")

# Some probes might not map to gene symbols, handle those
genes_62322 <- data.frame(
  probe_id = probe_ids_62322,
  gene_symbol = gene_symbols_62322,
  stringsAsFactors = FALSE
)

# Check how many probes were mapped to gene symbols
sum(!is.na(genes_62322$gene_symbol))
sum(is.na(genes_62322$gene_symbol))

# Create a new expression matrix with gene symbols as row names
# For probes that map to the same gene, we'll keep the one with highest variance
expr_with_genes_62322 <- data.frame(combined_samples_62322_norm)
expr_with_genes_62322$gene_symbol <- gene_symbols_62322

# Remove probes without gene symbols
expr_with_genes_62322 <- expr_with_genes_62322[!is.na(expr_with_genes_62322$gene_symbol), ]

# For probes mapping to the same gene, keep the one with highest variance
expr_with_genes_62322$variance <- apply(expr_with_genes_62322[, 1:(ncol(expr_with_genes_62322)-2)], 1, var)
expr_with_genes_62322 <- expr_with_genes_62322[order(expr_with_genes_62322$gene_symbol, 
                                                    -expr_with_genes_62322$variance), ]
expr_with_genes_62322 <- expr_with_genes_62322[!duplicated(expr_with_genes_62322$gene_symbol), ]

# Create final expression matrix with gene symbols as row names
expr_gse62322_genes <- expr_with_genes_62322[, 1:(ncol(expr_with_genes_62322)-2)]
rownames(expr_gse62322_genes) <- expr_with_genes_62322$gene_symbol
```

## GSE68204
```{r, include=FALSE}
# For GSE68204 (miRNA dataset), we need a different annotation approach
# The platform is GPL10850 (Agilent Human miRNA Microarray)
# Check the column containing miRNA names in the feature data
fData_68204 <- fData(gse68204[[1]])
head(fData_68204)

# Look for column that contains miRNA identifiers
# This might be "NAME", "SPOT_ID", or something else
miRNA_col <- grep("NAME|SPOT_ID|miRNA|ID", colnames(fData_68204), value = TRUE)[1]
miRNA_ids <- fData_68204[, miRNA_col]

# Create a mapping between probe IDs and miRNA names
miRNA_mapping <- data.frame(
  probe_id = rownames(fData_68204),
  miRNA_id = miRNA_ids,
  stringsAsFactors = FALSE
)

# Map expression data to miRNA IDs
expr_with_miRNA_68204 <- data.frame(expr_gse68204)
expr_with_miRNA_68204$miRNA_id <- miRNA_mapping$miRNA_id[match(rownames(expr_with_miRNA_68204), 
                                                               miRNA_mapping$probe_id)]
expr_with_miRNA_68204 <- expr_with_miRNA_68204[!is.na(expr_with_miRNA_68204$miRNA_id), ]

# For probes mapping to the same miRNA, keep the one with highest variance
expr_with_miRNA_68204$variance <- apply(expr_with_miRNA_68204[, 1:(ncol(expr_with_miRNA_68204)-2)], 1, var)
expr_with_miRNA_68204 <- expr_with_miRNA_68204[order(expr_with_miRNA_68204$miRNA_id, 
                                                    -expr_with_miRNA_68204$variance), ]
expr_with_miRNA_68204 <- expr_with_miRNA_68204[!duplicated(expr_with_miRNA_68204$miRNA_id), ]

# Create final expression matrix
expr_gse68204_miRNA <- expr_with_miRNA_68204[, 1:(ncol(expr_with_miRNA_68204)-2)]
rownames(expr_gse68204_miRNA) <- expr_with_miRNA_68204$miRNA_id
```

# DEG Identification
## GSE41258
```{r}
# GSE41258 (primary dataset)
design_41258 <- model.matrix(~0+factor(sample_type_41258))
colnames(design_41258) <- c("normal", "tumor")

# Fit the linear model
fit_41258 <- lmFit(expr_gse41258_genes, design_41258)

# Define the contrast matrix for tumor vs normal
contrast.matrix_41258 <- makeContrasts(tumor-normal, levels=design_41258)

# Fit the contrast
fit2_41258 <- contrasts.fit(fit_41258, contrast.matrix_41258)
fit2_41258 <- eBayes(fit2_41258)

# Get the results with adjusted p-values
results_41258 <- topTable(fit2_41258, number=Inf, adjust.method="BH")

# Filter DEGs based on criteria from the paper: adj.P.Val < 0.05 and |logFC| > 1
degs_41258 <- results_41258[results_41258$adj.P.Val < 0.05 & abs(results_41258$logFC) > 1, ]

# Check how many up and down-regulated genes
up_regulated_41258 <- degs_41258[degs_41258$logFC > 1, ]
down_regulated_41258 <- degs_41258[degs_41258$logFC < -1, ]

cat("Number of DEGs in GSE41258:", nrow(degs_41258), "\n")
cat("Up-regulated:", nrow(up_regulated_41258), "\n")
cat("Down-regulated:", nrow(down_regulated_41258), "\n")
```

## GSE62322
```{r}
# And for the validation dataset (GSE62322)
design_62322 <- model.matrix(~0+factor(sample_type_62322))
colnames(design_62322) <- c("normal", "tumor")

fit_62322 <- lmFit(expr_gse62322_genes, design_62322)
contrast.matrix_62322 <- makeContrasts(tumor-normal, levels=design_62322)
fit2_62322 <- contrasts.fit(fit_62322, contrast.matrix_62322)
fit2_62322 <- eBayes(fit2_62322)

results_62322 <- topTable(fit2_62322, number=Inf, adjust.method="BH")
degs_62322 <- results_62322[results_62322$adj.P.Val < 0.05 & abs(results_62322$logFC) > 1, ]

up_regulated_62322 <- degs_62322[degs_62322$logFC > 1, ]
down_regulated_62322 <- degs_62322[degs_62322$logFC < -1, ]

cat("Number of DEGs in GSE62322:", nrow(degs_62322), "\n")
cat("Up-regulated:", nrow(up_regulated_62322), "\n")
cat("Down-regulated:", nrow(down_regulated_62322), "\n")
```

## GSE68204
```{r}
# Repeat the same process for the miRNA dataset (GSE68204)
design_68204 <- model.matrix(~0+factor(sample_type_68204))
colnames(design_68204) <- c("normal", "tumor")

fit_68204 <- lmFit(combined_samples_68204_norm, design_68204)
contrast.matrix_68204 <- makeContrasts(tumor-normal, levels=design_68204)
fit2_68204 <- contrasts.fit(fit_68204, contrast.matrix_68204)
fit2_68204 <- eBayes(fit2_68204)

results_68204 <- topTable(fit2_68204, number=Inf, adjust.method="BH")
demirs_68204 <- results_68204[results_68204$adj.P.Val < 0.05 & abs(results_68204$logFC) > 1, ]

up_regulated_68204 <- demirs_68204[demirs_68204$logFC > 1, ]
down_regulated_68204 <- demirs_68204[demirs_68204$logFC < -1, ]

cat("Number of DEmiRs in GSE68204:", nrow(demirs_68204), "\n")
cat("Up-regulated:", nrow(up_regulated_68204), "\n")
cat("Down-regulated:", nrow(down_regulated_68204), "\n")
```

## Volcano Plots
```{r}
# Create volcano plots to visualize DEGs
# GSE41258
volcano_41258 <- data.frame(
  Gene = rownames(results_41258),
  logFC = results_41258$logFC,
  pvalue = -log10(results_41258$adj.P.Val)
)

volcano_41258$Significance <- ifelse(results_41258$adj.P.Val < 0.05 & abs(results_41258$logFC) > 1,
                                  ifelse(results_41258$logFC > 1, "Up-regulated", "Down-regulated"),
                                  "Not significant")

ggplot(volcano_41258, aes(x = logFC, y = pvalue, color = Significance)) +
  geom_point(alpha = 0.6) +
  scale_color_manual(values = c("Not significant" = "grey", 
                                "Up-regulated" = "red", 
                                "Down-regulated" = "blue")) +
  theme_minimal() +
  labs(title = "Volcano Plot of DEGs in GSE41258",
       x = "Log2 Fold Change",
       y = "-log10(adjusted p-value)") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  geom_vline(xintercept = c(-1, 1), linetype = "dashed")

# Save DEGs for WGCNA analysis
#write.csv(degs_41258, "degs_gse41258.csv")
#write.csv(demirs_68204, "demirs_gse68204.csv")
#write.csv(degs_62322, "degs_gse62322.csv")
```

##Clean Up Environment
```{r}
keep <- c("expr_gse41258_genes", "expr_gse62322_genes", "combined_samples_68204_norm", "degs_41258", "degs_62322", "degs_68204")
rm(list=setdiff(ls(), keep))
```

# WGCNA Analysis
## Prepare Data for WGCNA
```{r}
# Extract expression data for the DEGs from normalized data
degs_names <- rownames(degs_41258)
expr_data_degs <- expr_gse41258_genes[degs_names, ]

# Transpose the data for WGCNA (genes should be in columns)
expr_data_degs_t <- t(expr_data_degs)
```

## Soft-thresholding power (beta)
```{r}
# The paper mentioned they used a power of 6
powers <- c(1:20)
sft <- pickSoftThreshold(expr_data_degs_t, powerVector = powers, verbose = 5)

# First, create the jpg file
#jpeg(filename = "scale_free_topology_plots.jpg", width = 1000, height = 500, quality = 100)

# Plot the scale-free topology fit index as a function of the soft-thresholding power
par(mfrow = c(1, 2))
# Scale-free topology fit index as a function of the soft-thresholding power
plot(sft$fitIndices[, 1], -sign(sft$fitIndices[, 3]) * sft$fitIndices[, 2],
     xlab = "Soft Threshold (power)",
     ylab = "Scale Free Topology Model Fit, signed R^2",
     type = "n",
     main = paste("Scale independence"))
text(sft$fitIndices[, 1], -sign(sft$fitIndices[, 3]) * sft$fitIndices[, 2],
     labels = powers, cex = 0.9, col = "red")
abline(h = 0.90, col = "red")

# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[, 1], sft$fitIndices[, 5],
     xlab = "Soft Threshold (power)",
     ylab = "Mean Connectivity",
     type = "n",
     main = paste("Mean connectivity"))
text(sft$fitIndices[, 1], sft$fitIndices[, 5], labels = powers, cex = 0.9, col = "red")

#dev.off()
```

## Construct the co-expression network
```{r}
#power from the paper (power = 6)
softPower <- 14
adjacency <- adjacency(expr_data_degs_t, power = softPower)

# Step 4: Turn adjacency into topological overlap
TOM <- TOMsimilarity(adjacency)
dissTOM <- 1 - TOM

# Step 5: Cluster using TOM dissimilarity
geneTree <- hclust(as.dist(dissTOM), method = "average")

# Step 6: Define modules
# The paper doesn't specify the exact parameters, but we'll use standard ones
minModuleSize <- 30  # Minimum module size
# Module identification using dynamic tree cut
dynamicMods <- cutreeDynamic(dendro = geneTree, distM = dissTOM, 
                            deepSplit = 2, pamRespectsDendro = FALSE,
                            minClusterSize = minModuleSize)
table(dynamicMods)

# Convert numeric labels to colors
dynamicColors <- labels2colors(dynamicMods)
table(dynamicColors)

# Step 7: Merge modules whose expression profiles are very similar
# Calculate eigengenes
MEList <- moduleEigengenes(expr_data_degs_t, colors = dynamicColors)
MEs <- MEList$eigengenes

# Calculate dissimilarity of module eigengenes
MEDiss <- 1 - cor(MEs)
METree <- hclust(as.dist(MEDiss), method = "average")

# Plot the result
plot(METree, main = "Clustering of module eigengenes",
     xlab = "", sub = "")

# Set a cut height for merging modules (e.g., 0.25, meaning correlation of 0.75)
MEDissThres <- 0.25
merge <- mergeCloseModules(expr_data_degs_t, dynamicColors, cutHeight = MEDissThres, verbose = 3)
mergedColors <- merge$colors
mergedMEs <- merge$newMEs

# Plot the gene dendrogram and module colors after merging
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),
                   c("Dynamic Tree Cut", "Merged dynamic"),
                   dendroLabels = FALSE, hang = 0.03,
                   addGuide = TRUE, guideHang = 0.05)

# Save the module colors and merged module colors
moduleColors <- mergedColors
colorOrder <- c("grey", standardColors(50))
moduleLabels <- match(moduleColors, colorOrder) - 1
MEs <- mergedMEs
```

# Relate Modules to Clinical Traits
```{r}
# Extract tumor and normal samples
tumor_pheno_41258 <- pheno_gse41258[tumor_indices, ]
normal_pheno_41258 <- pheno_gse41258[normal_indices, ]

# Combine for DEG analysis later
combined_pheno_41258 <- rbind(tumor_pheno_41258, normal_pheno_41258)

# Extract clinical traits from the phenotype data based on column names
clinical_traits <- data.frame(row.names = colnames(combined_samples_41258))

# Extract age (numeric)
if ("age:ch1" %in% colnames(pheno_gse41258)) {
  clinical_traits$age <- as.numeric(pheno_gse41258[match(colnames(combined_samples_41258), rownames(pheno_gse41258)), "age:ch1"])
}

# Extract gender (M/F) and convert to binary (M=1, F=0)
if ("gender:ch1" %in% colnames(pheno_gse41258)) {
  clinical_traits$sex <- pheno_gse41258[match(colnames(combined_samples_41258), rownames(pheno_gse41258)), "gender:ch1"]
  # Convert to binary for correlation analysis
  clinical_traits$sex.binary <- ifelse(clinical_traits$sex == "M", 1, 0)
}

# Extract T stage (formatted as t: #)
if ("characteristics_ch1.7" %in% colnames(pheno_gse41258)) {
  t_stage_raw <- pheno_gse41258[match(colnames(combined_samples_41258), rownames(pheno_gse41258)), "characteristics_ch1.7"]
  # Extract number after "t: "
  clinical_traits$stage.t <- as.numeric(gsub("t: ", "", t_stage_raw))
}

# Extract N stage (formatted as n: #)
if ("characteristics_ch1.8" %in% colnames(pheno_gse41258)) {
  n_stage_raw <- pheno_gse41258[match(colnames(combined_samples_41258), rownames(pheno_gse41258)), "characteristics_ch1.8"]
  # Extract number after "n: "
  clinical_traits$stage.n <- as.numeric(gsub("n: ", "", n_stage_raw))
}

# Extract M stage (formatted as m: #)
if ("characteristics_ch1.9" %in% colnames(pheno_gse41258)) {
  m_stage_raw <- pheno_gse41258[match(colnames(combined_samples_41258), rownames(pheno_gse41258)), "characteristics_ch1.9"]
  # Extract number after "m: "
  clinical_traits$stage.m <- as.numeric(gsub("m: ", "", m_stage_raw))
}

# Extract overall stage (Roman numerals)
if ("group stage:ch1" %in% colnames(pheno_gse41258)) {
  stage_raw <- pheno_gse41258[match(colnames(combined_samples_41258), rownames(pheno_gse41258)), "group stage:ch1"]
  # Convert Roman numerals to numeric
  clinical_traits$stage <- stage_raw
  clinical_traits$stage.tot <- sapply(stage_raw, function(x) {
    if (is.na(x)) return(NA)
    switch(x,
           "I" = 1,
           "II" = 2,
           "III" = 3,
           "IV" = 4,
           NA)
  })
}

# Extract recurrence (Yes/No/NA)
if ("recurrence:ch1" %in% colnames(pheno_gse41258)) {
  clinical_traits$recurrence <- pheno_gse41258[match(colnames(combined_samples_41258), rownames(pheno_gse41258)), "recurrence:ch1"]
  # Convert to binary for correlation (Yes = 1, No = 0, NA = NA)
  clinical_traits$recurrence.binary <- sapply(clinical_traits$recurrence, function(x) {
    if (is.na(x)) return(NA)
    ifelse(x == "Yes", 1, 0)
  })
}

# Convert factor types for categorical variables
clinical_traits$sex <- factor(clinical_traits$sex)
clinical_traits$recurrence <- factor(clinical_traits$recurrence)

# Now relate modules to clinical traits
# Use all available numeric traits for correlation, including the binary sex variable
numeric_traits <- clinical_traits[, c("age", "sex.binary", "stage.t", "stage.n", "stage.m", "stage.tot", "recurrence.binary")]
numeric_traits <- numeric_traits[, sapply(numeric_traits, function(x) !all(is.na(x)))]

# Calculate the correlation between MEs and clinical traits
moduleTrait_cor <- cor(MEs, numeric_traits, use = "pairwise.complete.obs")
moduleTrait_pvalue <- corPvalueStudent(moduleTrait_cor, nrow(numeric_traits))

# Create a heatmap of module-trait relationships
textMatrix <- paste(signif(moduleTrait_cor, 2), "\n(", signif(moduleTrait_pvalue, 1), ")", sep = "")
dim(textMatrix) <- dim(moduleTrait_cor)

#jpeg(filename = "module_clinical_relationship.jpg", width = 1000, height = 500, quality = 100)

# Plot the heatmap
par(mar = c(6, 8.5, 3, 3))
labeledHeatmap(Matrix = moduleTrait_cor,
               xLabels = names(numeric_traits),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1, 1),
               main = paste("Module-trait relationships"))

#dev.off()
```
Go with turquoise, brown, or blue

# Analyze Turquoise Module
## Extract the Turquoise Module Genes
```{r}
# Extract the turquoise module genes
module_of_interest <- "turquoise"
turquoise_module_genes <- colnames(expr_data_degs_t)[which(moduleColors == module_of_interest)]

write(turquoise_module_genes, file = "turquoise_genes.txt")
```

## Functional Enrichment Analysis 
### Perform Analysis
```{r}
# Convert gene symbols to Entrez IDs for functional analysis
entrez_ids <- bitr(turquoise_module_genes, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)

# Perform GO enrichment analysis
go_bp <- enrichGO(gene = entrez_ids$ENTREZID, 
                  OrgDb = org.Hs.eg.db, 
                  ont = "BP", 
                  pAdjustMethod = "BH", 
                  pvalueCutoff = 0.05, 
                  qvalueCutoff = 0.05)

go_cc <- enrichGO(gene = entrez_ids$ENTREZID, 
                  OrgDb = org.Hs.eg.db, 
                  ont = "CC", 
                  pAdjustMethod = "BH", 
                  pvalueCutoff = 0.05, 
                  qvalueCutoff = 0.05)

go_mf <- enrichGO(gene = entrez_ids$ENTREZID, 
                  OrgDb = org.Hs.eg.db, 
                  ont = "MF", 
                  pAdjustMethod = "BH", 
                  pvalueCutoff = 0.05, 
                  qvalueCutoff = 0.05)

# Perform KEGG pathway analysis
kegg_enrich <- enrichKEGG(gene = entrez_ids$ENTREZID,
                          organism = "hsa",
                          pvalueCutoff = 0.05,
                          pAdjustMethod = "BH")
```

### Plot Results
```{r}
# Extract data from your enrichment results
# Get top 10 BP terms
bp_data <- go_bp@result %>%
  top_n(10, -p.adjust) %>%
  mutate(Category = "BP")

# Get top 10 CC terms
cc_data <- go_cc@result %>%
  top_n(10, -p.adjust) %>%
  mutate(Category = "CC")

# Get top 10 MF terms
mf_data <- go_mf@result %>%
  top_n(10, -p.adjust) %>%
  mutate(Category = "MF")

# Get top 5 KEGG terms
kegg_data <- kegg_enrich@result %>%
  top_n(5, -p.adjust) %>%
  mutate(Category = "KEGG")

# Combine all data
combined_data <- rbind(
  select(bp_data, Category, Term = Description, AdjPValue = p.adjust),
  select(cc_data, Category, Term = Description, AdjPValue = p.adjust),
  select(mf_data, Category, Term = Description, AdjPValue = p.adjust),
  select(kegg_data, Category, Term = Description, AdjPValue = p.adjust)
)

# Calculate -log10(p-value) for plotting
combined_data$NegLogPval <- -log10(combined_data$AdjPValue)

# Set the order of categories
combined_data$Category <- factor(combined_data$Category, levels = c("BP", "CC", "MF", "KEGG"))

# Create a function to order terms by -log10(p-value) within each category
create_ordered_terms <- function(data) {
  data %>%
    group_by(Category) %>%
    arrange(Category, desc(NegLogPval)) %>%
    ungroup() %>%
    mutate(Term = factor(Term, levels = Term[order(Category, desc(NegLogPval))]))
}

# Apply the ordering function
combined_data <- create_ordered_terms(combined_data)

# Set colors for different categories (matching the example image)
category_colors <- c("BP" = "#4DAF4A", "CC" = "#E41A1C", "MF" = "#984EA3", "KEGG" = "#377EB8")

# Create the plot
p_turq <- ggplot(combined_data, aes(x = NegLogPval, y = Term, fill = Category)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = category_colors) +
  labs(
    x = "-Log10(adj-P-value)",
    y = ""
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "left",
    legend.title = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(face = "bold")
  ) +
  facet_grid(Category ~ ., scales = "free_y", space = "free_y") +
  # Remove panel spacing to make it look more like the example
  theme(panel.spacing = unit(0, "lines"))

# Display the plot
print(p_turq)

# Save the plot
#ggsave("enrichment_barplot_grouped.png", p, width = 10, height = 10, dpi = 300)
```

# Analyze Blue Module
## Extract Blue Genes
```{r}
# Extract the blue module genes
module_of_interest <- "blue"
blue_module_genes <- colnames(expr_data_degs_t)[which(moduleColors == module_of_interest)]

write(blue_module_genes, file = "blue_genes.txt")
```

## Functional Enrichment Analysis 
### Perform Analysis
```{r}
# Convert gene symbols to Entrez IDs for functional analysis
entrez_ids <- bitr(blue_module_genes, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)

# Perform GO enrichment analysis
go_bp <- enrichGO(gene = entrez_ids$ENTREZID, 
                  OrgDb = org.Hs.eg.db, 
                  ont = "BP", 
                  pAdjustMethod = "BH", 
                  pvalueCutoff = 0.05, 
                  qvalueCutoff = 0.05)

go_cc <- enrichGO(gene = entrez_ids$ENTREZID, 
                  OrgDb = org.Hs.eg.db, 
                  ont = "CC", 
                  pAdjustMethod = "BH", 
                  pvalueCutoff = 0.05, 
                  qvalueCutoff = 0.05)

go_mf <- enrichGO(gene = entrez_ids$ENTREZID, 
                  OrgDb = org.Hs.eg.db, 
                  ont = "MF", 
                  pAdjustMethod = "BH", 
                  pvalueCutoff = 0.05, 
                  qvalueCutoff = 0.05)

# Perform KEGG pathway analysis
kegg_enrich <- enrichKEGG(gene = entrez_ids$ENTREZID,
                          organism = "hsa",
                          pvalueCutoff = 0.05,
                          pAdjustMethod = "BH")
```

### Plot Results
```{r}
# Extract data from your enrichment results
# Get top 10 BP terms
bp_data <- go_bp@result %>%
  top_n(10, -p.adjust) %>%
  mutate(Category = "BP")

# Get top 10 CC terms
cc_data <- go_cc@result %>%
  top_n(10, -p.adjust) %>%
  mutate(Category = "CC")

# Get top 10 MF terms
mf_data <- go_mf@result %>%
  top_n(10, -p.adjust) %>%
  mutate(Category = "MF")

# Get top 5 KEGG terms
kegg_data <- kegg_enrich@result %>%
  top_n(5, -p.adjust) %>%
  mutate(Category = "KEGG")

# Combine all data
# Now use dplyr's select with the explicit namespace
combined_data <- rbind(
  dplyr::select(bp_data, Category, Term = Description, AdjPValue = p.adjust),
  dplyr::select(cc_data, Category, Term = Description, AdjPValue = p.adjust),
  dplyr::select(mf_data, Category, Term = Description, AdjPValue = p.adjust),
  dplyr::select(kegg_data, Category, Term = Description, AdjPValue = p.adjust)
)

# Calculate -log10(p-value) for plotting
combined_data$NegLogPval <- -log10(combined_data$AdjPValue)

# Set the order of categories
combined_data$Category <- factor(combined_data$Category, levels = c("BP", "CC", "MF", "KEGG"))

# Create a function to order terms by -log10(p-value) within each category
create_ordered_terms <- function(data) {
  data %>%
    group_by(Category) %>%
    arrange(Category, desc(NegLogPval)) %>%
    ungroup() %>%
    mutate(Term = factor(Term, levels = Term[order(Category, desc(NegLogPval))]))
}

# Apply the ordering function
combined_data <- create_ordered_terms(combined_data)

# Set colors for different categories (matching the example image)
category_colors <- c("BP" = "#4DAF4A", "CC" = "#E41A1C", "MF" = "#984EA3", "KEGG" = "#377EB8")

# Create the plot
p_blue <- ggplot(combined_data, aes(x = NegLogPval, y = Term, fill = Category)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = category_colors) +
  labs(
    x = "-Log10(adj-P-value)",
    y = ""
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "left",
    legend.title = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(face = "bold")
  ) +
  facet_grid(Category ~ ., scales = "free_y", space = "free_y") +
  # Remove panel spacing to make it look more like the example
  theme(panel.spacing = unit(0, "lines"))

# Display the plot
print(p_blue)

# Save the plot
ggsave("blue_enrichment_barplot_grouped.png", p_blue, width = 10, height = 10, dpi = 300)
```

## PPI Analysis
### Cytoscape Hub Genes
Use the previously exported file, blue_genes.txt, and upload this to STRINGdb. Set minimum required interaction score to 0.700 and hide disconnected nodes in the network. From there, send the network to cytoscape. Using the cytoHubba app, find the top 10 genes using MCC, MNC, and DEGREE. 
```{r}
blue_mcc <- c("COL1A2", "COL10A1", "SERPINH1", "COL1A1", "COL5A2", "COL11A1", "COL5A1", "BGN", "COL2A1", "THBS2")
blue_mnc <- c("SPP1", "COL1A2", "COL10A1", "COL1A1", "COL5A2", "TIMP1", "COL11A1", "COL5A1", "BGN", "THBS2")
blue_degree <- c("SPP1", "COL1A2", "COL10A1", "COL1A1", "COL5A2", "TIMP1", "COL11A1", "COL5A1", "BGN", "THBS2")

#Get intersection of these lists for later
blue_top <- Reduce(intersect, list(blue_mcc, blue_mnc, blue_degree))

# Create the Venn diagram using ggVennDiagram
ggvenn <- ggVennDiagram(
  list(
    MCC = blue_mcc,
    MNC = blue_mnc,
    Degree = blue_degree
  ),
  label = "count",
  category.names = c("MCC", "MNC", "Degree")
)

# Customize the appearance
ggvenn <- ggvenn +
  scale_fill_gradient(low = "white", high = "skyblue") +
  theme(legend.position = "none")

# Display the plot
print(ggvenn)
```

#### Volcano Plot Labeled w/ Hub Genes
```{r}
ggplot(volcano_41258, aes(x = logFC, y = pvalue, color = Significance)) +
  geom_point(alpha = 0.6) +
  scale_color_manual(values = c("Not significant" = "grey", 
                                "Up-regulated" = "red", 
                                "Down-regulated" = "blue")) +
  theme_minimal() +
  labs(title = "Volcano Plot of DEGs in GSE41258",
       x = "Log2 Fold Change",
       y = "-log10(adjusted p-value)") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  geom_vline(xintercept = c(-1, 1), linetype = "dashed") +
  # Add repelled labels with arrows and black text
  geom_label_repel(data = genes_to_label, 
                   aes(label = Gene),
                   fill = "white",
                   color = "black",         # Black text color
                   segment.color = "black", # Black arrow color
                   arrow = arrow(length = unit(0.01, "npc")), # Add arrows
                   box.padding = 0.5,
                   point.padding = 0.3,
                   force = 2,
                   max.overlaps = 30,
                   size = 3)
```


### Plot STRING
```{r}
# Connect to STRING database
string_db <- STRINGdb$new(version="12.0", species=9606, score_threshold=700)

# Map your turquoise genes to STRING identifiers
mapped_genes <- string_db$map(data.frame(gene=blue_module_genes), "gene", removeUnmappedRows = TRUE)

# Get interactions between these genes
interactions <- string_db$get_interactions(mapped_genes$STRING_id)

# Create mapping between STRING IDs and gene symbols
id_mapping <- mapped_genes[, c("STRING_id", "gene")]
id_to_symbol <- setNames(id_mapping$gene, id_mapping$STRING_id)

# Create an igraph object
g <- graph_from_data_frame(interactions[, c("from", "to", "combined_score")], directed = FALSE)

# Get largest connected component if you want to focus on main network
components <- components(g)
largest_component <- which.max(components$csize)
subg <- induced_subgraph(g, which(components$membership == largest_component))

# Get edges for the subgraph
subg_edges <- as_data_frame(subg)

# Function to pad gene names with spaces to match the longest name
pad_gene_name <- function(names) {
  # Find the longest name
  max_length <- max(nchar(names))
  
  # Pad each name with spaces to match the longest name
  padded_names <- sapply(names, function(name) {
    spaces_needed <- max_length - nchar(name)
    left_spaces <- floor(spaces_needed / 2)
    right_spaces <- ceiling(spaces_needed / 2)
    return(paste0(
      paste(rep("  ", left_spaces), collapse = ""),
      name,
      paste(rep("  ", right_spaces), collapse = "")
    ))
  })
  
  return(padded_names)
}

# Get all the gene symbols
gene_symbols <- sapply(V(subg)$name, function(x) ifelse(!is.na(id_to_symbol[x]), id_to_symbol[x], x))
gene_symbols <- unname(gene_symbols)

# Apply padding to make all labels the same length
padded_symbols <- pad_gene_name(gene_symbols)

# Calculate node degree for sizing
node_degrees <- degree(subg)

# Define the list of genes you want to highlight
highlighted_genes <- blue_top

# Create nodes dataframe with custom coloring using separate columns
nodes <- data.frame(
  id = V(subg)$name,
  label = padded_symbols,
  title = gene_symbols,
  shape = "circle",
  size = 40,
  stringsAsFactors = FALSE
)

# Define separate color columns
nodes$color.background <- ifelse(gene_symbols %in% highlighted_genes, "#FF9999", "#D3D3D3")
nodes$color.border <- ifelse(gene_symbols %in% highlighted_genes, "#CC0000", "#A9A9A9")

# Correctly create the font column: one list per node
nodes$font <- replicate(nrow(nodes), list(color = "black", size = 40, face = "normal"), simplify = FALSE)

# Create edges dataframe (unchanged)
edges <- data.frame(
  from = subg_edges$from,
  to = subg_edges$to,
  value = 10,
  title = paste("Confidence:", subg_edges$combined_score),
  color = "lightblue",
  arrows = "false",
  stringsAsFactors = FALSE
)

# Create the network visualization with more spread-out nodes
network <- visNetwork(nodes, edges, width = "2000px", height = "1500px") %>%
  visNodes(scaling = NULL) %>%
  visPhysics(
    solver = "forceAtlas2Based", 
    forceAtlas2Based = list(
      gravitationalConstant = -800,
      centralGravity = 0.01,
      springLength = 200,
      springConstant = 0.05,
      damping = 0.4,
      avoidOverlap = 1
    ),
    stabilization = list(
      enabled = TRUE, 
      iterations = 2000,
      updateInterval = 25
    )
  ) %>%
  visEdges(smooth = FALSE) %>%
  visLegend(width = 0.1, position = "right", main = "Turquoise Gene Network") %>%
  visInteraction(dragNodes = TRUE, dragView = TRUE, zoomView = TRUE) %>%
  visEvents(
    stabilized = "function() { this.setOptions({physics: false}); }"
  ) %>%
  visLayout(randomSeed = 123)

# Render the network in RStudio
network

# Save the network as an HTML file
#saveWidget(network, "turquoise_gene_network.html", selfcontained = TRUE)

# Install phantomjs if not already installed
# webshot::install_phantomjs()

# Save a high-resolution screenshot
#webshot("turquoise_gene_network.html", "turquoise_gene_network.png", 
#        delay = 3, vwidth = 1750, vheight = 1500)
```


# Analyze Brown Module
```{r}
# Extract the turquoise module genes
module_of_interest <- "brown"
brown_module_genes <- colnames(expr_data_degs_t)[which(moduleColors == module_of_interest)]

write(brown_module_genes, file = "brown_genes.txt")
```

## Functional Enrichment Analysis 
### Perform Analysis
```{r}
# Convert gene symbols to Entrez IDs for functional analysis
entrez_ids <- bitr(turquoise_module_genes, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)

# Perform GO enrichment analysis
go_bp <- enrichGO(gene = entrez_ids$ENTREZID, 
                  OrgDb = org.Hs.eg.db, 
                  ont = "BP", 
                  pAdjustMethod = "BH", 
                  pvalueCutoff = 0.05, 
                  qvalueCutoff = 0.05)

go_cc <- enrichGO(gene = entrez_ids$ENTREZID, 
                  OrgDb = org.Hs.eg.db, 
                  ont = "CC", 
                  pAdjustMethod = "BH", 
                  pvalueCutoff = 0.05, 
                  qvalueCutoff = 0.05)

go_mf <- enrichGO(gene = entrez_ids$ENTREZID, 
                  OrgDb = org.Hs.eg.db, 
                  ont = "MF", 
                  pAdjustMethod = "BH", 
                  pvalueCutoff = 0.05, 
                  qvalueCutoff = 0.05)

# Perform KEGG pathway analysis
kegg_enrich <- enrichKEGG(gene = entrez_ids$ENTREZID,
                          organism = "hsa",
                          pvalueCutoff = 0.05,
                          pAdjustMethod = "BH")
```

### Plot Results
```{r}
# Extract data from your enrichment results
# Get top 10 BP terms
bp_data <- go_bp@result %>%
  top_n(10, -p.adjust) %>%
  mutate(Category = "BP")

# Get top 10 CC terms
cc_data <- go_cc@result %>%
  top_n(10, -p.adjust) %>%
  mutate(Category = "CC")

# Get top 10 MF terms
mf_data <- go_mf@result %>%
  top_n(10, -p.adjust) %>%
  mutate(Category = "MF")

# Get top 5 KEGG terms
kegg_data <- kegg_enrich@result %>%
  top_n(5, -p.adjust) %>%
  mutate(Category = "KEGG")

# Combine all data
combined_data <- rbind(
  select(bp_data, Category, Term = Description, AdjPValue = p.adjust),
  select(cc_data, Category, Term = Description, AdjPValue = p.adjust),
  select(mf_data, Category, Term = Description, AdjPValue = p.adjust),
  select(kegg_data, Category, Term = Description, AdjPValue = p.adjust)
)

# Calculate -log10(p-value) for plotting
combined_data$NegLogPval <- -log10(combined_data$AdjPValue)

# Set the order of categories
combined_data$Category <- factor(combined_data$Category, levels = c("BP", "CC", "MF", "KEGG"))

# Create a function to order terms by -log10(p-value) within each category
create_ordered_terms <- function(data) {
  data %>%
    group_by(Category) %>%
    arrange(Category, desc(NegLogPval)) %>%
    ungroup() %>%
    mutate(Term = factor(Term, levels = Term[order(Category, desc(NegLogPval))]))
}

# Apply the ordering function
combined_data <- create_ordered_terms(combined_data)

# Set colors for different categories (matching the example image)
category_colors <- c("BP" = "#4DAF4A", "CC" = "#E41A1C", "MF" = "#984EA3", "KEGG" = "#377EB8")

# Create the plot
p_brown <- ggplot(combined_data, aes(x = NegLogPval, y = Term, fill = Category)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = category_colors) +
  labs(
    x = "-Log10(adj-P-value)",
    y = ""
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "left",
    legend.title = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(face = "bold")
  ) +
  facet_grid(Category ~ ., scales = "free_y", space = "free_y") +
  # Remove panel spacing to make it look more like the example
  theme(panel.spacing = unit(0, "lines"))

# Display the plot
print(p_brown)

# Save the plot
#ggsave("enrichment_barplot_grouped.png", p, width = 10, height = 10, dpi = 300)
```

# Validate Hub Genes
## GSE62322
```{r}
# Replace this list with your actual hub genes from Cytoscape analysis
hub_genes <- blue_top
# Check if these genes are in the validation dataset
valid_hub_genes <- hub_genes[hub_genes %in% rownames(expr_gse62322_genes)]
if (length(valid_hub_genes) > 0) {
  # Extract expression values for the hub genes
  hub_gene_expr <- expr_gse62322_genes[valid_hub_genes, ]
  
  # Calculate mean expression in tumor vs normal
  tumor_indices <- (length(normal_indices_62322)+1):(length(normal_indices_62322)+length(tumor_indices_62322))
  normal_indices <- 1:length(normal_indices_62322)
  
  # Calculate mean expression in tumor vs normal
  tumor_mean <- rowMeans(hub_gene_expr[, tumor_indices])
  normal_mean <- rowMeans(hub_gene_expr[, normal_indices])
  
  # Calculate log2 fold change
  log2fc <- tumor_mean - normal_mean
  
  # Calculate p-values
  pvalues <- numeric(length(valid_hub_genes))
  for (i in 1:length(valid_hub_genes)) {
    gene <- valid_hub_genes[i]
    t_test <- t.test(hub_gene_expr[gene, tumor_indices],
                     hub_gene_expr[gene, normal_indices])
    pvalues[i] <- t_test$p.value
  }
  
  # Create a dataframe for plotting
  hub_validation <- data.frame(
    gene = valid_hub_genes,
    normal_mean = normal_mean,
    tumor_mean = tumor_mean,
    log2fc = log2fc,
    pvalue = pvalues
  )
  
  # Add significance stars
  hub_validation$stars <- ""
  hub_validation$stars[hub_validation$pvalue < 0.05] <- "*"
  hub_validation$stars[hub_validation$pvalue < 0.01] <- "**"
  hub_validation$stars[hub_validation$pvalue < 0.001] <- "***"
  
  # Add significance levels for legend
  hub_validation$significance <- "ns"
  hub_validation$significance[hub_validation$pvalue < 0.05] <- "p < 0.05"
  hub_validation$significance[hub_validation$pvalue < 0.01] <- "p < 0.01"
  hub_validation$significance[hub_validation$pvalue < 0.001] <- "p < 0.001"
  
  # Order by log2fc for better visualization
  hub_validation <- hub_validation[order(hub_validation$log2fc), ]
  
  # Order the genes by log2 fold change for better visualization
  hub_validation$gene <- factor(hub_validation$gene, levels = hub_validation$gene[order(hub_validation$log2fc)])
  
  # Create a new column for bar colors (red for positive, blue for negative log2fc)
  hub_validation$color <- ifelse(hub_validation$log2fc > 0, "red", "blue")
  
  # Create a separate data frame for the legend
  legend_data <- data.frame(
    label = c("p < 0.05", "p < 0.01", "p < 0.001", "Not significant"),
    symbol = c("*", "**", "***", "")
  )
  
  # Build the plot
  p <- ggplot(hub_validation, aes(x = gene, y = log2fc, fill = color)) +
    geom_bar(stat = "identity", width = 0.7) +
    coord_flip() +
    geom_text(aes(label = stars, 
                  y = ifelse(log2fc > 0, log2fc + 0.2, log2fc - 0.2)),
              color = "black", size = 5) +
    scale_fill_identity() +  # Uses the color values directly
    labs(title = "Hub Gene Expression in GSE62322",
         x = "Gene",
         y = "Log2 Fold Change (Tumor vs Normal)") +
    theme_minimal(base_size = 14) +
    theme(plot.title = element_text(hjust = 0.5),
          axis.text.y = element_text(face = "bold", color = "black"),
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank(),
          legend.position = "bottom") +
    # Add manual legend for significance
    annotate("text", x = Inf, y = Inf, 
             label = "Significance:
             * p < 0.05   
            ** p < 0.01   
                 *** p < 0.001", 
             hjust = 1.2, vjust = 4, size = 4)
  
  # Add a horizontal line at y=0
  p <- p + geom_hline(yintercept = 0, linetype = "dashed", color = "gray50")
  
  # Print the plot
  print(p)
  
} else {
  cat("None of the hub genes found in the validation dataset.\n")
}
```

## TGCA-COAD
```{r}
# Your hub genes from blue module
blue_top

# Query and download TCGA-COAD data
query_coad <- GDCquery(
  project = "TCGA-COAD", 
  data.category = "Transcriptome Profiling",
  data.type = "Gene Expression Quantification",
  workflow.type = "STAR - Counts"
)

# Download the data (only run this once - it's time-consuming)
#GDCdownload(query_coad)

# Prepare the data
coad_data <- GDCprepare(query_coad)

# Extract the count data
count_matrix <- assay(coad_data)

# Extract sample type information to identify tumor and normal samples
sample_info <- colData(coad_data)
tumor_samples <- which(sample_info$sample_type %in% c("Primary Tumor"))
normal_samples <- which(sample_info$sample_type %in% c("Solid Tissue Normal"))

# Normalize the data (using log2 transformation of TPM)
normalized_counts <- log2(count_matrix + 1)

# Function to get gene expression by gene name
get_gene_expr <- function(gene_name) {
  gene_id <- rowData(coad_data)[which(rowData(coad_data)$gene_name == gene_name), "gene_id"]
  if(length(gene_id) == 0) {
    return(NULL)
  }
  return(normalized_counts[gene_id, ])
}

# Extract expression values for the hub genes
hub_gene_expr <- data.frame()
for(gene in blue_top) {
  expr <- get_gene_expr(gene)
  if(!is.null(expr)) {
    tumor_expr <- expr[tumor_samples]
    normal_expr <- expr[normal_samples]
    
    # Calculate mean expression and p-value
    tumor_mean <- mean(tumor_expr, na.rm = TRUE)
    normal_mean <- mean(normal_expr, na.rm = TRUE)
    log2fc <- tumor_mean - normal_mean
    
    t_test <- t.test(tumor_expr, normal_expr)
    pvalue <- t_test$p.value
    
    # Add to dataframe
    hub_gene_expr <- rbind(hub_gene_expr, 
                           data.frame(Gene = gene, 
                                      Normal_mean = normal_mean,
                                      Tumor_mean = tumor_mean,
                                      Log2FC = log2fc,
                                      Pvalue = pvalue))
  }
}

# Add significance stars
hub_gene_expr$Stars <- ""
hub_gene_expr$Stars[hub_gene_expr$Pvalue < 0.05] <- "*"
hub_gene_expr$Stars[hub_gene_expr$Pvalue < 0.01] <- "**"
hub_gene_expr$Stars[hub_gene_expr$Pvalue < 0.001] <- "***"

# Order by log2FC for better visualization
hub_gene_expr <- hub_gene_expr[order(hub_gene_expr$Log2FC), ]
hub_gene_expr$Gene <- factor(hub_gene_expr$Gene, levels = hub_gene_expr$Gene)

# Create bar plot similar to Figure 11 in the paper
tcga_validation_plot <- ggplot(hub_gene_expr, aes(x = Gene, y = Log2FC, fill = ifelse(Log2FC > 0, "Up", "Down"))) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Stars, y = ifelse(Log2FC > 0, Log2FC + 0.2, Log2FC - 0.2)),
            size = 5) +
  scale_fill_manual(values = c("Up" = "red", "Down" = "blue"),
                    name = "Regulation") +
  labs(title = "Validation of Hub Genes in TCGA-COAD",
       x = "Gene",
       y = "Log2 Fold Change (Tumor vs Normal)") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank()
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50")

print(tcga_validation_plot)
```

# Survival Analysis
This was done at the GEPIA2 website (http://gepia2.cancer-pku.cn/#survival)
```{r}
# Directory containing your PDF plots
pdf_dir <- "Survival_Analysis"

# Get list of PDF files
pdf_files <- list.files(pdf_dir, pattern = "*.pdf", full.names = TRUE)

# Function to convert PDF to a raster object that can be plotted
pdf_to_grob <- function(pdf_file) {
  # Convert PDF to bitmap
  pdf_bitmap <- pdftools::pdf_convert(pdf_file, dpi = 300, format = "png", pages = 1)
  
  # Read the PNG file
  img <- png::readPNG(pdf_bitmap)
  
  # Create a raster grob
  grob <- rasterGrob(img, interpolate = TRUE)
  
  # Clean up temporary PNG file
  file.remove(pdf_bitmap)
  
  return(grob)
}

# Convert all PDFs to grobs
plot_grobs <- lapply(pdf_files, pdf_to_grob)

# Create the 4x2 grid layout
pdf("combined_survival_plots.pdf", width = 15, height = 7)
grid.arrange(grobs = plot_grobs, 
             ncol = 4, 
             nrow = 2,
             heights = c(1, 1),  # Equal heights for both rows
             padding = unit(0.1, "line"))  # Reduce padding between plots
dev.off()

# You can also save as PNG if needed
png("combined_survival_plots.png", width = 3000, height = 1500, res = 300)
grid.arrange(grobs = plot_grobs, ncol = 4, nrow = 2)
dev.off()
```


# Bipartite miRNA-mRNA Network Analysis
```{r}
# Get the list of differentially expressed miRNAs
demir_names <- rownames(demirs_68204)

# Format miRNA names to standard format if needed
# multiMiR expects names like "hsa-miR-940"
# Check if they are in the correct format
if (!all(grepl("^hsa-", demir_names))) {
  # Preprocess miRNA names to ensure they have the "hsa-" prefix
  demir_names <- sapply(demir_names, function(x) {
    if (!grepl("^hsa-", x)) {
      return(paste0("hsa-", x))
    } else {
      return(x)
    }
  })
}

# Get predicted targets for the DEmiRs (this can take some time)
mir_targets <- get_multimir(org = "hsa", 
                           mirna = demir_names,
                           target = blue_module_genes,
                           table = "validated",
                           summary = TRUE)

# Extract the results
mir_target_df <- mir_targets@data

# Save the miRNA-mRNA interactions
#write.csv(mir_target_df, "mir_target_interactions.csv")
```

```{r}
# Filter for interactions involving hub genes (based on the paper)
hub_genes <- blue_top
hub_gene_targets <- mir_target_df[mir_target_df$target_symbol %in% hub_genes, ]
hub_gene_targets <- hub_gene_targets[, c(3, 4)]
colnames(hub_gene_targets) <- c("Mature miRNA ID", "Gene Target")

# Create a publication-ready table without row names
kbl <- kable(hub_gene_targets,
             row.names = FALSE,                     # Remove row names
             booktabs = TRUE, 
             align = "c") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                latex_options = c("hold_position", "scale_down"),
                full_width = FALSE)

kbl
```

## Survival Analysis of miRNAs
```{r}
#hsa_miR_125a_5p.csv
data <- read.csv("miRNA_survival/hsa_miR_125a_5p.csv", header = TRUE, stringsAsFactors = FALSE)
data$status_numeric <- ifelse(data$Status == "Dead", 1, 0)

# Fit the Kaplan-Meier survival model
# Replace 'time', 'status', and 'group' with your actual column names if they differ
fit <- survfit(Surv(Days, status_numeric) ~ Group, data = data)

# Plot the survival curve with confidence intervals and a risk table
plot1 <- ggsurvplot(
  fit, 
  data = data,
  pval = TRUE,
  conf.int = FALSE,
  risk.table = FALSE,
  title = "KaplanMeier Survival Curve for hsa_miR_125a_5p",
  xlab = "Time (months)",
  ylab = "Survival Probability",
  palette = c("blue", "red")
)

#hsa_miR_133b.csv
data <- read.csv("miRNA_survival/hsa_miR_133b.csv", header = TRUE, stringsAsFactors = FALSE)
data$status_numeric <- ifelse(data$Status == "Dead", 1, 0)

# Fit the Kaplan-Meier survival model
# Replace 'time', 'status', and 'group' with your actual column names if they differ
fit <- survfit(Surv(Days, status_numeric) ~ Group, data = data)

# Plot the survival curve with confidence intervals and a risk table
plot2 <- ggsurvplot(
  fit, 
  data = data,
  pval = TRUE,
  conf.int = FALSE,
  risk.table = FALSE,
  title = "KaplanMeier Survival Curve for hsa_miR_133b",
  xlab = "Time (months)",
  ylab = "Survival Probability",
  palette = c("blue", "red")
)

#hsa_miR_532_5p.csv
data <- read.csv("miRNA_survival/hsa_miR_532_5p.csv", header = TRUE, stringsAsFactors = FALSE)
data$status_numeric <- ifelse(data$Status == "Dead", 1, 0)

# Fit the Kaplan-Meier survival model
# Replace 'time', 'status', and 'group' with your actual column names if they differ
fit <- survfit(Surv(Days, status_numeric) ~ Group, data = data)

# Plot the survival curve with confidence intervals and a risk table
plot3 <- ggsurvplot(
  fit, 
  data = data,
  pval = TRUE,
  conf.int = FALSE,
  risk.table = FALSE,
  title = "KaplanMeier Survival Curve for hsa_miR_532_5p",
  xlab = "Time (months)",
  ylab = "Survival Probability",
  palette = c("blue", "red")
)

#hsa_miR_940.csv
data <- read.csv("miRNA_survival/hsa_miR_940.csv", header = TRUE, stringsAsFactors = FALSE)
data$status_numeric <- ifelse(data$Status == "Dead", 1, 0)

# Fit the Kaplan-Meier survival model
# Replace 'time', 'status', and 'group' with your actual column names if they differ
fit <- survfit(Surv(Days, status_numeric) ~ Group, data = data)

# Plot the survival curve with confidence intervals and a risk table
plot4 <- ggsurvplot(
  fit, 
  data = data,
  pval = TRUE,
  conf.int = FALSE,
  risk.table = FALSE,
  title = "KaplanMeier Survival Curve for hsa_miR_940",
  xlab = "Time (months)",
  ylab = "Survival Probability",
  palette = c("blue", "red")
)

arranged_plots <- arrange_ggsurvplots(list(plot1, plot2, plot3, plot4), 
                                      ncol = 2, nrow = 2,
                                      print = TRUE)
```

